%----------------------------------------------------------------------------------------------------
% START OF CHAPTER
%----------------------------------------------------------------------------------------------------
\chapter{May 4}
\paragraph{Topics}
\textit{Direct and indirect object pronouns, Sentence review}

\section{Grammar} \label{0504-Grammar}
\subsection{Direct object pronouns}

A direct object pronoun replaces the noun object in a sentence. In English this corresponds
to `it', `he/she', or `them'. I eat \textit{it}. He saw \textit{her}. They stand in for
whatever thing or person the verb is acting on.

Note the subtle distinction between who/what the verb is acting \textit{on} and
who/what the verb is acting \textit{for}. Direct object pronouns are for the person or
thing that the verb is acting \textit{on}.

\begin{mdframed}
	\textbf{Direct object pronouns} in Italian replace nouns to avoid repetition.
	They answer the questions `what?' or `whom?' in relation to the verb.
	The pronoun goes before the conjugated verb.
\end{mdframed}

\begin{table}[ht]
	\centering
	\begin{tabular}{ r | l }
	                         & pronoun   \\
		\hline
		\textbf{me}          & \iti{mi}  \\
		\textbf{you}         & \iti{ti}  \\
		\textbf{him/it}      & \iti{lo}  \\
		\textbf{her/it}      & \iti{la}  \\
		\textbf{us}          & \iti{ci}  \\
		\textbf{you pl.}     & \iti{vi}  \\
		\textbf{them, masc.} & \iti{li}  \\
		\textbf{them, fem.}  & \iti{le}
	\end{tabular}
	\caption{Direct object pronouns}
	\label{tab:tDirectObjectPronouns}
\end{table}

In the past (perfect) tense, the past participle must be modified, even with \iti{avere},
to match the gender/plurality of the pronoun, as if it is working as an adjective.
Also, the \iti{lo, li} forms contract to \iti{l'} before the \textit{h}, e.g.
\iti{La macchina? Sì, l´ho vista.}

\paragraph{Examples}
\begin{itemize}
	\item{I eat gelato. $\rightarrow$ I eat it.}       \\
	\iti{Mangio gelato.} $\rightarrow$ \iti{\uline{Lo} mangio.}

	\item{I love pizza. I eat pizza every day. $\rightarrow$ I love pizza. I eat it every day.} \\
	\iti{Amo la pizza. Mangio la pizza ogni giorno.} $\rightarrow$ \iti{Amo la pizza. \uline{La} mangio ogni giorno.}

	\item{They saw me at the restaurant.} \\
	\iti{\uline{Mi} hanno visto al ristorante.}

	\item{She saw the paintings in the museum. $\rightarrow$ She saw them in the museum.} \\
	\iti{Ha visto i quadri nel museo $\rightarrow$ \uline{Li} ha visto nel museo.}

	\item{I bought the car. I bought it.} \\
	\iti{Ho comprato la macchina. L'ho comprata.}

	\item{Ho visto che tu l'hai comprato già.} \\
	\iti{I saw that you had already bought it. (masc.)}

	\item {Ho visto che tu l'hai comprata già.} \\
	\iti{I saw that you had already bought it. (fem.)}
\end{itemize}

\subsection{Indirect object pronouns}
These pronouns are used to indicate who or what is receiving the effect of the verb.
I wrote \textit{him} a letter. He gave \textit{her} the book.

Look for verbs like \iti{dare, scrivere, telefonare, dire, parlare, chiedere} when considering
if an indirect object pronoun is appropriate, since all these verbs have some kind of target (recipient):
to give (someone) something, to write (to someone), to phone (someone), to tell (someone),
to speak (to someone), to ask (someone).

\begin{mdframed}
	\textbf{Indirect object pronouns} indicate the recipient or beneficiary of the action,
	answering `to whom?' or `for whom?' the action is done. If you could say, `I [verb] to [name] something',
	then [name] can become an indirect pronoun.
\end{mdframed}

\begin{table}[ht]
	\centering
	\begin{tabular}{ r l | l }
	               & English           & Italian    \\
		\hline
			to/for & \textbf{me}       & \iti{mi}   \\
			       & \textbf{you}      & \iti{ti}   \\
                   & \textbf{him}      & \iti{gli}  \\
                   & \textbf{her}      & \iti{le}   \\
                   & \textbf{us}       & \iti{ci}   \\
                   & \textbf{you pl.}  & \iti{vi}   \\
                   & \textbf{them}     & \iti{gli}
	\end{tabular}
	\caption{Indirect object pronouns}
	\label{tab:tIndirectObjectPronouns}
\end{table}

Tip: if in doubt, decide what the verb is acting upon (the direct object), and
to whom or for whom the action is done (the indirect object).

\paragraph{Examples}
\begin{itemize}
	\item{I told you that I bought the ice cream for tonight.} \\
	\iti{Ti ho detto che ho comprato il gelato per questa sera.}

	\item{I told you already that I worked a lot last night.} \\
	\iti{Ti ho detto già che ho lavorato molto ieri sera.}

    \item{I'm writing a letter to Marco. $\rightarrow$ I'm writing him a letter.} \\
    \iti{Scrivo una lettera a Marco. $\rightarrow$ Gli scrivo una lettera.}

    \item{I gave the book to Elena. $\rightarrow$ I gave the book to her.} \\
    \iti{Ho dato il libro a Elena. $\rightarrow$ Le ho dato il libro.}
\end{itemize}

\section{Sentence review} \label{0504-Sentences}
\begin{itemize}
	\item{I'm happy because my wife already bought a table for our living room. We needed this table
		for when we have guests for dinner. The table is big and is for twelve people.
		I'm happy that she has already bought it.} \\
	\iti{Sono felice perché mia moglie ha già comprato un tavolo per il nostro soggiorno.
		Abbiamo bisogno di questo tavolo quando abbiamo ospiti per cena. Il tavolo è grande
		ed è per dodici persone. Sono felice che lo ha comprato già.}

	\item{Hi Giorgio! I called you last night but you didn't answer. I hope everything is OK.
		I'm working a lot these weeks. I told you that I wanted to go on vacation with you and your family.
		Let's talk about that soon. See you soon!} \\
	\iti{Ciao Giorgio! Ho ti chiamato ieri sera, ma tu non hai risposto. Spero tutto bene. Sto lavorando
		molto queste settimane. Ti ho detto che voglio andare in vacanza con te e la tua famiglia.
		Parliamo di questo presto. A presto!}
\end{itemize}

%----------------------------------------------------------------------------------------------------
% END OF CHAPTER
%----------------------------------------------------------------------------------------------------